---
title: "R Stats Functions"
author: "Jonah Davids"
date: "01/06/2020"
output: html_document
---

https://homerhanumat.github.io/tigerstats

```{r load-packages, message=FALSE}
library(statsr)
library(dplyr)
library(shiny)
library(ggplot2)
library(tigerstats)
library(GGally)
```

***Normal distribution***


Lets starts with the pnorm() function. **(Use this if you have the point estimate and want to know the p-value)**

Don't forget that for two sided tests, we often multiply the final p value by 2. 

The function pnorm() in regular R, as well as the function pnormGC() in thetigerstats` package, compute probabilities from known bounding values. For example, suppose that X is a normally distributed random variable with mean 70 and standard deviation 3, and that you want to know:

P(X<72). [the probability that a given value is lower than 72]

Then you know the boundary value 72, but you don’t know the probability: the area under the normal density curve before 72. Functions like pnormGC() aim to give you that area–that probability:

```{r}
pnormGC(72, region="below", mean=70, sd=3,graph=TRUE)
        
```
We can also use the pnorm() function to do something similar. 

```{r}
pnorm(72, mean=70, sd=3, lower.tail = TRUE)
```

This gives us the same value, but without the explanatory graph.

Both of these functions look at the area below a given value. However, we can analyze the area above the value with a simple adjustment from region "below" to "above" or lower.tail "TRUE" to "FALSE". 

```{r}
pnormGC(72, region="above", mean=70, sd=3,graph=TRUE)
```

with pnorm()

```{r}
pnorm(72, mean=70, sd=3, lower.tail = FALSE)
```

It is also possible for us to find how much area falls in between two point estimates using the pnormGC() function.

```{r}
pnormGC(c(-1.96,1.96),region="between",mean=0, sd=1,graph=TRUE)
```


Now, let's take a look at the qnorm function. **Use this if you have the centile/p-value and want to know the point estimate**

qnorm() does the opposite of pnorm. Given an area, it finds the boundary value that determines the area.

For example, suppose you want to find that 85th percentile of a normal distribution whose mean is 70 and whose standard deviation is 3. Then you ask for:

```{r}
qnorm(0.85,mean=70,sd=3, lower.tail = TRUE)
```

The value 73.1093 is indeed the 85th percentile, in the sense that 85% of the values in a population that is normally distributed with mean 70 and standard deviation 3 will lie below 73.1093. In other words, if you were to pick a random member X
 from such a population, then

P(X<73.1093) = 0.85

You can check that this is correct by plugging 73.1093 into pnormGC():

```{r}
pnormGC(73.1093,region="below",mean=70, sd=3,graph=TRUE)
```

***T distribution***

Lets start with the pt() function. **This is for when you have the point estimate and want the p-value.**

This works similarly to the pnorm function described above. Lets say I have a point estimate of 10 with 50 degrees of freedom:

```{r}
pt(1.69, df = 30, lower.tail = TRUE)
```

This prints 0.9493024, about 0.95%, which is the corresponding p-value. 

Now lets examine the qt() function. **This is for when you have the centile/p-value and want to know the point estimate**

the qt() function works similarly to the qnorm() function. Both have you input a centile/p-value and get out your point estimate. Lets take the above p-value of 0.9493024, still with df = 30.

```{r}
qt(0.9493024, df = 30, lower.tail = TRUE)
```

This prints out 1.69, which lo and behold, was the point estimate we used to get 0.9493024 in the first place.


***F distribution***

Lets look at the pf() function. **This is for when you already have a point estimate and are looking for the critical value.**

Lets say I have a point estimate of 21.735. My first sample has a df = 3, and the second sample has a df = 731.

```{r}
pf(21.735, 3, 791, lower.tail = FALSE)
```

Now, lets check out the qf() function. **This is for when you already have critical value and are looking for the point estimate.**

Lets use the above critical value of 1.559855e-13, still with df = 3 and df = 791).

```{r}
qf(1.559855e-13, 3, 791, lower.tail = FALSE)
```

There we go!

*** Proportion***

If we want to find the proportion of a statistic, we can use the prop.test() function, inputting our sample statistic and the N.

```{r}
prop.test(2.36, 200)
```

***Inference Tests***

This allows us to do hypothesis tests (T test).


```{r}
inference(y = weight, x = Garage.Yr.Built, data = nc, statistic = "mean", type = "ht", null = 0, 
          alternative = "twosided", method = "theoretical")
```

We can also use this function to look for confidence intervals. 

```{r}
inference(y = weeks, data = nc, statistic = "mean", type = "ci", conf_level = 0.95, null = 0, 
          alternative = "twosided", method = "theoretical")
```

Calculating proportions are also a possibility:

```{r}
inference(y = response, data = us12, statistic = "proportion", type = "ci", method = "theoretical", success = "atheist")
```




***Chi-square statistics***

Lets say we have one categorical variable with multiple levels. Having computed the test statistic, we can use the pchisq() function to find its corresponding p-value.

```{r}
pchisq(22.63, 4, lower.tail = FALSE)
```

For a better visualiztion, we can also use the pchisqGC function from tigerstats.

```{r}
pchisqGC(22.63, region = "above", df = 4, graph = TRUE)
```

***Correlation***

If we simply want to know what R is, we can use the summarise feature like so:

```{r}
mlb11 %>%
  summarise(cor(runs, at_bats))
```

Alternatively, we can quickly graph this and get a through line using ss_plot.

```{r}
plot_ss(x = at_bats, y = runs, data = mlb11)
```

If we want to find statistics for the regression line, we can use the 1m function and then summary.

```{r}
m1 <- lm(runs ~ at_bats, data = mlb11)

summary(m1)
```

If we want to find a specific y given a specific x, we can use filter and select to do the following:

```{r}
mlb11 %>%
  filter(at_bats == 5579) %>%
  select(runs)
```

**Scatterplot correlation with regression line**

```{r}
ggplot(data = evals, aes(x = bty_avg, y = score)) +
  geom_jitter() +
  geom_smooth(method = "lm")
```


**Residual normal plot (for checking whether residuals are normal)**

```{r}
qqnorm(resid(model1))
```

Examining relationships bwtween many many variables at once with ggpairs:

```{r}
library(GGally)

ggpairs(evals, columns = 13:19)
```

**Prediction intervals**

If we want to predict a specific item using a linear model, we can first create that predicted model with a new data frame and then use the predict function. 

```{r}
newprof <- data.frame(gender = "male", bty_avg = 3)

predict(m_bty_gen, newprof, interval = "prediction", level = 0.95)
```

